### 2019-09-06 peer-review/review-response.Rmd

I also found it harder than it should have been to understand exactly what was different between this analysis and the original analysis, and which differences mattered most (which is important as it is the main aim of the work as I understand it). I wonder if a table would be useful detailing the differences: any implementation/modelling differences; differences in point estimates of d (both from methods and use of more recent data?); differences in parameters from literature; inclusion of uncertainty in each. I think it's important to differentiate between differences that could have been included in the analysis at the time (eg uncertainty) vs things that have emerged since (eg updated evidence on serial intervals). I also think it's important to be clear whether methods, central estimates, or uncertainty matter most from the point of view of decision-relevant changes in the evidence generated.

*To answer your previous comment we introduced a subsection of the model explanation detailing our additions to the model. The Updating model parameter estimates and incorporating parameter uncertainty section of the methods contains a discussion of differences between the original and updated analysis. We have reworded this for clarity (please see the marked up paper). As you say, parameter uncertainty for incidence rates could have been included at the time but was not, whilst the updated value for the serial interval has only recently become available. This is now explicitly stated in the reworded methods paragraph.*

*We agree that it is important to be clear which findings are relevant to policy makers and in what context. We have rewored the 4th discussion paragraph as follows in order to provide more context.*

*"This study indicates that some of the evidence used to justify the 2005 change in BCG vaccination policy may have underestimated the impact of ending the scheme. It highlights the importance of including both parameter and measurement error, as excluding these sources of variation may lead to spuriously precise results. This is problematic for policy makers as the worst case scenario often needs to be considered when making policy decisions. In addition, our exploration of the assumptions used to estimate the annual change in TB incidence rates in the ethnic White UK born illustrates the structural impact of assuming an annual decrease in TB incidence rates. More realistic estimates of the annual decrease in incidence rates resulted in a greatly increased impact of ending the BCG schools scheme. Policy makers should consider these updated estimates when assessing the role of BCG vaccination in those at school-age. However, decisions regarding vaccine policy require economic evaluation, which discounts costs and benefits in the future; discounting has not been applied in this study which estimates the epidemiological impact of vaccination only."*

@@

*We agree that these estimates (from the original Sutherland et al. paper) are very unrealistic. Sutherland et al. assumed a year-on-year decline of 9% with an estimated incidence of  ~ 23 per 100,000 in 1983 for the unvaccinated population. In 2013 this would give an incidence rate of ~ 2.6. These very small incidence rates mean that the difference between incidence rates in unvaccinated and vaccinated cases is also very small leading to the spuriously high number need to vaccinate. In comparison a 2% decline would give an inidence rate of ~ 12.5. Our additions to the methods section add extra detail on how these estimates are arrived at.*

- I think discounting needs to be discussed at some point. This is almost always included in decision analyses.

*We agree that discounting should be included for economic evaluation. However we have not seen it done widely when the study is focusing purely on the epidemiological impact of vaccination. We have added the following sentences to the discussion,*

*"Policy makers should consider these updated estimates when assessing the role of BCG vaccination in those at school-age. However, decisions regarding vaccine policy require economic evaluation, which discounts costs and benefits in the future; discounting has not been applied in this study which estimates the epidemiological impact of vaccination only."*

### 2019-10-04 peer-review/review-response.Rmd

10.	Page 8, Line 40-41: ‘A scenario with a 1.9% annual decrease in incidence rates was most comparable to our results based on notifications.’ As the medium annual decrease estimated using notifications was 3.13% (Page 7, Line 20), why the scenario of a 3.9% annual decrease did not give the most comparable results?

*Unfortunately, the summary statistics do not capture changes in the annual decrease in notifications over time. Notifications first increased and then decreased and it is these trends that led to the overall 1.9% decrease being a better fit. To help clarify this we have moved a figure from the SI into the main body of the paper.*

11.	Page 9, Line 37-38: Do you consider to report the primary impacts of ending the BCG programme and the additional cases due to transmission separately? This will show the importance of including transmission structure in assessing the BCG vaccination policy.

*We have separated the reporting of primary and secondary additional notifications to highlight the role of the transmission model.*

@@

### OVERVIEW

This paper revisits an analysis that was used to support the decision to discontinue the UK policy of BCG vaccination in schools. It considers updated evidence on parameters as well as incorporating uncertainty in model inputs. Re-evaluating modelling around decisions and considering the impact of uncertainty are both under-explored and worthy topics.

At the moment, I think this article needs major revisions before publication as it is extremely difficult to follow exactly what has been done.

*Thank you for your comments - especially the identification of the flaw with the original Sutherland et al. model. Based on them we have corrected the transmission chain model, extended our discussion of the methods used, and added further clarification where possible. Please see our detailed responses below.*

@@

### MAJOR ISSUES

The idea of the approach is that information about the rate of decline of TB incidence over time combined with knowledge of the generation time should inform on the net reproduction number, and hence the scope for indirect benefits from preventing cases. However, I really can't follow the model description on page 5. Part 1: T (and in part 2) is described as the sum over all generations of resulting cases, but R0^z (presumably R rather than R0) and (1-d)^z seem very much like the expected size of the z-th generation of resulting cases? Part 2: x is described as an initial size, but it seems in fact to be a net reproduction number from the way it is used in a geometric progression (resulting cases should scale linearly with initial size under branching process type models). If x=R and eqn 1 is true, things seem over-determined: one can solve R(z) without need for data on d?? Part 3 makes sense if x is R. Then I also don't understand how this relates to the 3 equations at the bottom of the page. I also don't remember the eligibility for BCG vaccination being introduced (line 37), and am not clear what it means to 'reproduce the distribution of cases due to transmission over time'. I think all this needs much more careful explanation.

*We agree that this needed further explanation and that the definition used in 1. of Sutherland et al. was incorrect. We have corrected the Sutherland model and extended the model explanation.*

@@

I also found it harder than it should have been to understand exactly what was different between this analysis and the original analysis, and which differences mattered most (which is important as it is the main aim of the work as I understand it). I wonder if a table would be useful detailing the differences: any implementation/modelling differences; differences in point estimates of d (both from methods and use of more recent data?); differences in parameters from literature; inclusion of uncertainty in each. I think it's important to differentiate between differences that could have been included in the analysis at the time (eg uncertainty) vs things that have emerged since (eg updated evidence on serial intervals). I also think it's important to be clear whether methods, central estimates, or uncertainty matter most from the point of view of decision-relevant changes in the evidence generated.

*To answer your previous comment we introduced a subsection of the model explanation detailing our additions to the model. The Updating model parameter estimates and incorporating parameter uncertainty section of the methods contains a discussion of differences between the original and updated analysis. We have reworded this for clarity (please see the marked up paper). As you say, parameter uncertainty for incidence rates could have been included at the time but was not, whilst the updated value for the serial interval has only recently become available. This is now explicitly stated in the reworded methods paragraph.*

*We agree that it is important to be clear which findings are relevant to policy makers and in what context. We have rewored the 4th discussion paragraph to provide more context.*

@@

*Thank you - corrected throughout.*

- In a couple of places (ie strengths/limitations bullets & Discussion) it is stated this is not "the most accurate method" for assessing the impact of these changes. This does beg questions about what other methods would have been better, and in what respects (eg how is "accuracy" defined?). This is worth discussing. I think it also means that the rationale for the work is about assessing the imapct of updated/modified analysis for inputs within an existing framework, rather than trying to provide a better answer to the same question (otherwise, why use a suboptimal method?)

*We agree that this could have been better phrased. We have rephrased as follows,*

*" The simulation approach used here, although updated where possible, is not the most accurate method for estimating the impact of ending the BCG schools scheme as it relies on numerous assumptions based on the available knowledge in 1987 and does not account for the role of non-White and non-UK born cases. However, the strength of this work is that the estimates were based on the framework used to inform policy making. This allowed the strength of model used in the decision-making process to be assessed once parameter uncertainty had been incorporated and for flaws in the model to be identified. This would not have been possible if the impact had been assessed using only the observed data or with a alternative model. It also allowed estimates based on updated data to be compared to historic estimates within the same framework. This would also not have been possible if a different framework had been used."*

@@

*Thanks for pointing this out. Whilst there has been a systematic review the authors identifed only two papers with estimates (the one used here and one other older paper). Both papers made use of similar datasets in the Netherlands. The cited paper is 10 years newer, has an improved sample size, and improved methods.*

- pg 6 paragraph on methods to quantify decline: again, I found this confusing and it is a key part of the analysis. Also, what is "local regression" please?

*We have reworded this paragraph (and made changes suggested by reviewer 1). Please see the marked up paper. We have also included a graph in the results from the SI that shows  the changes over time in the annual decline depending on the data/assumptions used.*

*We have added the following clarification for local regression,*

*"LOESS (locally estimated scatterplot smoothing) is a local regression method. It combines multiple regression models in a k-nearest neighbours meta-model. This approach allows nonlinear trends to be fitted to using a series of linear models."*

@@

- pg 7 paragraph on incidence rate changes: many medians presented do not lie between the corresponding 2.5% and 97.5% quantiles, which must be typos. Also there seems to be some confusion about whether to present decreases or negative increases.

*Unfortunately, our choice of "-" to separate estimates of the 2.5% and 97.5% quantiles made this paragraph unclear. We have moved to using "," throughout the paper as this more clearly identifies when estimates are negative. Thank you for picking this up.*

@@

This manuscript reports the results of a mathematical modeling analysis that updates a previous analysis done in 1987 to project the potential effect of ending universal school-age vaccination in the UK. It uses the same model framework but incorporates uncertainty estimates (which the previous model did not) and explores the effect of changing one of the original model assumptions about rate of decline in TB incidence to what has actually been observed in the intervening decades.

*Thank you for your comments.*

I think that the authors have executed their stated goal well, but I am having a hard time seeing this as an independently publishable piece of work. Making minor tweaks to the model parameters and adding uncertainty estimates did not produce fundamentally new conclusions. The more important change in the assumption of rate of decline in TB incidence did change the results, but the conclusions still do not seem independently publishable; the conclusion seems to be simply that the projected impact depends very much on what you assume is happening with the epidemic. I wonder if maybe I am not understanding the results completely? It seems like since 15 years have passed since the policy change, the more useful analysis would be to try to assess what the impact has been in those 15 years rather than repeat the projection exercise that led to the policy change with varying assumptions about declines in TB incidence when in fact it should be knowable what has happened in those 15 years. Maybe this is what the authors were trying to do, but I find myself getting to the end of the article with the unanswered question about whether the policy change was the right decision or not.

*We agree that this study does not fully answer whether or not the policy change was the correct decision and agree that analysing recent data may be a more valuable approach for this (see here for a preprint by the authors doing this: https://doi.org/10.1101/567511). The aim of this study was instead to assess the strength of the evidence used for decision making at the time of the policy change - we have updated the abstract and introduction to clarify this. As mentioned by the second reviewer models are rarely evaluated once the policy decision has been made. In this study we show that the model used for decision making was methodologically flawed, was spuriously precise, and made several assumptions that led to what we now know were entirely unrealistic estimates. These findings are important for decision makers to understand when evaluating quantitative evidence for future policy decisions.*

### 2019-10-07 peer-review/review-response.Rmd

*Thank you for your comments - especially the identification of the flaw with the original Sutherland et al. model. Based on them we have corrected the transmission chain model, extended our discussion of the methods used, and added further clarification where possible. Please see our detailed responses below.*

### MAJOR ISSUES

The idea of the approach is that information about the rate of decline of TB incidence over time combined with knowledge of the generation time should inform on the net reproduction number, and hence the scope for indirect benefits from preventing cases. However, I really can't follow the model description on page 5. Part 1: T (and in part 2) is described as the sum over all generations of resulting cases, but R0^z (presumably R rather than R0) and (1-d)^z seem very much like the expected size of the z-th generation of resulting cases? Part 2: x is described as an initial size, but it seems in fact to be a net reproduction number from the way it is used in a geometric progression (resulting cases should scale linearly with initial size under branching process type models). If x=R and eqn 1 is true, things seem over-determined: one can solve R(z) without need for data on d?? Part 3 makes sense if x is R. Then I also don't understand how this relates to the 3 equations at the bottom of the page. I also don't remember the eligibility for BCG vaccination being introduced (line 37), and am not clear what it means to 'reproduce the distribution of cases due to transmission over time'. I think all this needs much more careful explanation.

*We agree that this needed further explanation and that the definition used in 1. of Sutherland et al. was incorrect. We have corrected the Sutherland model and extended the model explanation to include this.*

@@

*Thank you - corrected throughout.*

- In a couple of places (ie strengths/limitations bullets & Discussion) it is stated this is not "the most accurate method" for assessing the impact of these changes. This does beg questions about what other methods would have been better, and in what respects (eg how is "accuracy" defined?). This is worth discussing. I think it also means that the rationale for the work is about assessing the imapct of updated/modified analysis for inputs within an existing framework, rather than trying to provide a better answer to the same question (otherwise, why use a suboptimal method?)

*We agree that this could have been better phrased. We have rephrased as follows,*

*"The simulation approach used here, although updated where possible, is not the most accurate method for estimating the impact of ending the BCG schools scheme as it relies on numerous assumptions based on the available knowledge in 1987 and does not account for the role of non-White and non-UK born cases. However, the strength of this work is that the estimates were based on the framework used to inform policy making. This allowed the strength of model used in the decision-making process to be assessed once parameter uncertainty had been incorporated and for flaws in the model to be identified. This would not have been possible if the impact had been assessed using only the observed data or with a alternative model. It also allowed estimates based on updated data to be compared to historic estimates within the same framework. This would also not have been possible if a different framework had been used."*

### 2019-10-07 README.Rmd
editing and cuts to word length

## Abstract

### Objectives

@@

### Design

We recreated an approach for estimating the impact of ending the BCG schools scheme, updating the model with parameter uncertainty and improved parameter estimates where possible. We investigated scenarios for the assumed annual decrease in TB incidence rates considered by the UK’s Joint Committee on Vaccination and Immunisation and explored alternative scenarios using notification data.

### Setting

@@

### Outcome measures

@@

### Results

The model produced predictions with wide confidence intervals when parameter uncertainty was included. Estimates using the original assumptions greatly underpredicted the impact of ending school-age vaccination compared to estimates using notification data. Updating the assumed annual decrease in TB incidence rates resulted in estimates that were more comparable with notification data with an annual decrease of 1.9% representing the closest estimate to the data. Using this assumption, we estimate that 1600 (2.5,97.5% Quantiles (Q): 1300, 2100) vaccines would have been required to prevent a single notification in 2004.

### Conclusions

The previously published estimates of the impact of ending the BCG schools scheme were found to be spuriously precise once updated with parameter uncertainty. Our estimates of the impact of ending school-age vaccination were highly sensitive to the annual decrease in incidence rates. The impact of ending the BCG schools’ scheme was found to be greater than previously thought when notification data were used. Our results highlight the importance of including uncertainty and evaluating multiple scenarios when forecasting the impact of changes in vaccination policy.

>>>>

## Objectives

@@

## Design

We recreated an approach for estimating the impact of ending the BCG schools scheme, correcting a methodological flaw in the model, updating the model with parameter uncertainty, and improving parameter estimates where possible. We investigated scenarios for the assumed annual decrease in TB incidence rates considered by the UK’s Joint Committee on Vaccination and Immunisation and explored alternative scenarios using notification data.

## Setting

@@

## Outcome measures

@@

## Results

The original model greatly underestimated the impact of ending school-age vaccination compared to our updated, corrected, model. The model produced predictions with wide confidence intervals when parameter uncertainty was included. Estimates were most comparable to notification data when an annual decrease in TB incidence rates of 1.9% was used. Using this assumption, we estimate that 1600 (2.5,97.5% Quantiles (Q): 1300, 2000) vaccines would have been required to prevent a single notification in 2004.

## Conclusions

The previously published model was found to contain a methodological flaw which, once corrected, resulted in a large increase in the estimated impact of ending the BCG schools scheme. The original approach was also found to be spuriously precise once updated with parameter uncertainty. The impact of ending the BCG schools’ scheme was found to be greater than previously thought when notification data were used. Our results highlight the importance of independent evaluations of modelling evidence, including uncertainty, and evaluating multiple scenarios when forecasting the impact of changes in vaccination policy.
